---
title: "Activity validation"
output:
  html_document: default
date: "2025-12-21"
---




```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,warning=F,message=F,error=F,cache=F,fig.height = 5,fig.width = 11)

library(factoextra)
library(readxl)
library(multcomp)
library(caret)
library(tibble)
library(mgcv)
library(emmeans)
library(patchwork)
library(plotly)
library(ggplot2)
library(dendextend)
library(ggsci)
library(tidyr)
library(data.table)
library(dplyr)
library(ggbeeswarm)
library(ggprism)
library(sgpv)
library(robustlmm)
library(glmmTMB)
library(RColorBrewer)
library(see)
library(ggpattern)
library(MASS)
library(geomtextpath)



```

# Section {.tabset}



```{r, include=FALSE}

setwd("~/")


params = read_excel("activity_parameters_awaketoasleep.xlsx")

params = na.omit(params)
parms = tolower(params$...3)

file_path = "~/MiMIC"

MiMIC = TRUE
Trojan = FALSE

# Colors can be set here

if(MiMIC){
  custom_colors =   custom_colors = rep(c( "#737373","#e7cdb4" ),2)
}

if(Trojan){
  custom_colors = rep(c( "#737373", "#94c8c9"),2)
}


files = list.files(file_path,pattern = ".*.xlsx")

df=NULL
for(i in files){
  combine = NULL
  for(j in 1:length(excel_sheets( paste0(file_path,"/",i)))){
      tmp = read_excel( paste0(file_path,"/",i),sheet=j)
      colnames(tmp) = make.names(colnames(tmp))
      colnames(tmp)[which(substr(tolower(colnames(tmp)),1,3)=="fly")] = "Fly.Number"
      Measure = excel_sheets( paste0(file_path,"/",i))[j] 
      colnames(tmp)[5:ncol(tmp)] = paste0(Measure,colnames(tmp)[5:ncol(tmp)])
      tmp = tmp[,! colnames(tmp) %in% c("Bin.Length","Bin Length")]
      if(j == 1){
        combine = tmp
      }  else {
        combine = merge(combine,tmp,by=c("Group","Replicate","Fly.Number"))
      }
    }
  combine$file = i
  df = rbind(df,combine)
}

df = df[! df$Fly.Number %in% c( "Average" ,"SEM"), ]
df[is.na(df)] = 0
sds = apply(df[4:(ncol(df)-1)], 2, function(x) sd(x))
ix = which(!is.na(sds))

df = df[grepl("MGD",fixed=T,df$Group),]

colnames(df) = gsub("X","",colnames(df))

```


```{r,include=F}

fixed = gsub("X24","24",gsub("X12","12",gsub("Bin","",gsub("2nd","12h2",gsub("1st","12h1",gsub("BinAVG","",gsub("\\..*","",gsub("\\.bin","",gsub("\\.hour\\.bin","",colnames(df))))))))))

matched = sapply(parms, function(p) which(grepl(p ,fixed=T, tolower(fixed)) ))

#Hard coding for convenience, double check that 247 is file, 1:3 are group, replicate, and fly number

# df = df[,c(1:3,matched,247)]

sds = apply(df[4:(ncol(df)-1)], 2, function(x) sd(x))
ix = which(!is.na(sds))

```


```{r,fig.height = 6,fig.width=18}

df$Replicate = paste0(df$Group,df$Replicate,df$file)
df$Diet = sapply(strsplit(df$Group,"-"),function(x) x[[2]]) # leaving this logic in for future use
df$Genotype = sapply(strsplit(df$Group,"-"),function(x) x[[1]])
df$Sex = sapply(strsplit(df$Group,"-"),function(x) x[[3]])

convert = read_excel("screen_gene_names (2).xlsx")
# convert = read_excel("/Users/Ellen.Stirtz/OneDrive - Van Andel Institute/thesis/GxE/screen_gene_names.xlsx")


names = convert$`Drosophila Ortholog`[ match(unique(df$Genotype[df$Genotype != "WT"]),convert$`Bloomington Number` )]

names[is.na(names)] = unique(df$Genotype[df$Genotype != "WT"])[is.na(names)]

names = gsub("-",".",names)

df$Genotype = factor(df$Genotype,levels=c("WT",unique(as.character(df$Genotype[df$Genotype != "WT"]))),labels = c("WT", names))

# For every genotype and sex, 
# we fit a PCA
# apply rotation matrix to controls
# analyze controls for null interval
# analyze mutants to see if 95% CI does not overlap control effects

res = NULL

gns = unique(df$Genotype) 
gns = as.character(gns[gns != "WT" & ! grepl( "Ctrl",fixed=T, gns)] )

fixed2 = gsub("X24","24",gsub("X12","12",gsub("Bin","",gsub("2nd","12h2",gsub("1st","12h1",gsub("BinAVG","",gsub("\\..*","",gsub("\\.bin","",gsub("\\.hour\\.bin","",colnames(df))))))))))

```

```{r}

matched2 = sapply(parms, function(p) which(grepl(p ,fixed=T, tolower(fixed2)) ))
# matched2 = colnames(df)[4:(ncol(df)-4)]

```


```{r}
#Rose plot logic

# Light/Dark outcomes
outcomes_ld <- params$`...3`[
  grepl("light", params$Description, fixed = TRUE) |
    grepl("dark",  params$Description, fixed = TRUE)
]

# Start with the explicit 24h tags
outcomes_24_explicit <- params$`...3`[
  grepl("24h",     params$Description, fixed = TRUE) |
    grepl("24 hr",   params$Description, fixed = TRUE) |
    grepl("24-hour", params$Description, fixed = TRUE) |
    grepl("overall", params$Description, fixed = TRUE) |
    grepl("daily",   params$Description, fixed = TRUE)
]

# If you don't have explicit 24h tags, define 24h as "everything NOT LD"
if (length(outcomes_24_explicit) == 0) {
  outcomes_24 <- setdiff(params$`...3`, outcomes_ld)
} else {
  # If you do have explicit tags, STILL enforce exclusivity
  outcomes_24 <- setdiff(outcomes_24_explicit, outcomes_ld)
}

rose_plot <- function(res_indv, sex, diet, outcomes, mode = c("24h", "ld")) {
  mode <- match.arg(mode)
  
  new.names <- gsub("Bin","",gsub("2nd","12h2",gsub("1st","12h1",
                                                    gsub("BinAVG","",gsub("\\..*","",gsub("\\.bin","",
                                                                                          gsub("\\.hour\\.bin","",res_indv$Outcome)))))))
  
  match.indv <- unlist(sapply(tolower(new.names), function(p) which(grepl(p, fixed=TRUE, parms))))
  res_indv$names <- parms[match.indv]
  
  res_indv$light <- ifelse(grepl("12h1", fixed=TRUE, res_indv$names),"Light",
                           ifelse(grepl("12h2", fixed=TRUE, res_indv$names),"Dark","24h"))
  
  res_indv$Outcome <- res_indv$names
  
  plot_data <- subset(res_indv[res_indv$Sex == sex,], select = c(Outcome, estimate, p.value, light))
  plot_data <- plot_data[plot_data$Outcome %in% outcomes,]
  
  # MODE filter (this is the key feature you wanted)
  if (mode == "24h") {
    plot_data <- subset(plot_data, light == "24h")
  } else {
    plot_data <- subset(plot_data, light %in% c("Light","Dark"))
  }
  
  if (nrow(plot_data) == 0) {
    message("No rows for sex=", sex, " mode=", mode, " after filtering.")
    return(invisible(NULL))
  }
  
  plot_data$direction <- ifelse(plot_data$estimate < 0,"Down","Up")
  plot_data$ratio <- 2^abs(plot_data$estimate)
  plot_data$time_direction <- paste0(plot_data$direction," (",plot_data$light,")")
  plot_data$time_direction[plot_data$time_direction == "NA (NA)"] <- NA
  plot_data$ratio[plot_data$ratio > 3] <- 3
  
  plot_data$Outcome <- factor(plot_data$Outcome)
  plot_data$Outcome <- gsub("_.*","", params$`MORE ACTIVE`[match(plot_data$Outcome, params$...3)])
  lvls <- unique(plot_data$Outcome)
  
  plot_data <- plot_data[order(plot_data$Outcome, plot_data$light),]
  
  empty_bar <- 2
  to_add <- data.frame(matrix(NA, empty_bar*nlevels(plot_data$Outcome), ncol(plot_data)))
  colnames(to_add) <- colnames(plot_data)
  to_add$Outcome <- rep(levels(plot_data$Outcome), each=empty_bar)
  
  plot_data <- rbind(plot_data, to_add) %>% arrange(Outcome, light)
  
  label_data <- plot_data %>% group_by(Outcome) %>% summarize(tot=max(ratio,na.rm=TRUE), .groups="drop")
  number_of_bar <- nrow(label_data)
  angle <- 90 - 360 * (as.numeric(factor(label_data$Outcome,levels=lvls))-0.5) / number_of_bar
  label_data$hjust <- ifelse(angle < -90, 1, 0)
  label_data$angle <- ifelse(angle < -90, angle+180, angle)
  
  plot_data$Outcome <- factor(plot_data$Outcome, levels=lvls)
  
  base_data <- plot_data %>% group_by(Outcome) %>%
    summarize(start=min(as.numeric(as.factor(Outcome))), end=max(as.numeric(as.factor(Outcome))), .groups="drop") %>%
    rowwise() %>% mutate(title=mean(c(start,end)))
  
  grid_data <- base_data
  grid_data$end <- grid_data$end[c(nrow(grid_data), 1:nrow(grid_data)-1)] + 1
  grid_data$start <- grid_data$start - 1
  grid_data <- grid_data[-1,]
  
  if (plot_data$light[1] == "24h") {
    plot_data$time_direction <- factor(plot_data$time_direction, levels=c("Up (24h)","Down (24h)"))
  } else {
    plot_data$time_direction <- factor(plot_data$time_direction,
                                       levels=c("Up (Dark)","Down (Dark)","Up (Light)","Down (Light)"))
  }
  
  p.hold <- ggplot(plot_data, aes(x=as.factor(Outcome), y=ratio, fill=time_direction, color=time_direction)) +
    geom_bar(stat="identity", position=position_dodge(width=0.85), width=0.5) +
    geom_segment(data=grid_data, aes(x=end, y=2, xend=start, yend=2),
                 colour="gray", linewidth=0.3, inherit.aes=FALSE) +
    geom_segment(data=grid_data, aes(x=end, y=1, xend=start, yend=1),
                 colour="gray", linewidth=0.3, inherit.aes=FALSE) +
    geom_segment(data=grid_data, aes(x=end, y=0, xend=start, yend=0),
                 colour="gray", linewidth=0.3, inherit.aes=FALSE) +
    annotate("text", x=rep(min(as.numeric(plot_data$Outcome))-0.3,3), y=c(0,1,2),
             label=c("1","2","3+"), color="grey30", size=7, hjust=1) +
    coord_polar(start=0, clip="off") +
    theme_minimal(12) + ylim(c(-0.5,4)) +
    theme(legend.position="bottom", axis.text=element_blank(), axis.title=element_blank(),
          panel.grid=element_blank(), plot.margin=unit(rep(0,4),"cm")) +
    geom_text(data=label_data, aes(x=Outcome,label=Outcome,hjust=hjust,angle=angle),
              y=0.5, color="black", size=0, angle=label_data$angle, inherit.aes=FALSE)
  
  if (plot_data$light[1] == "24h") {
    p.hold <- p.hold +
      scale_color_manual(values=c("Down (24h)"="white","Up (24h)"="white"), na.translate=FALSE) +
      scale_fill_manual(values=c("Down (24h)"=alpha("darkblue",0.3),
                                 "Up (24h)"=alpha("darkred",0.3)), na.translate=FALSE) +
      theme(legend.title=element_blank())
  } else {
    p.hold <- p.hold +
      scale_color_manual(values=c("Down (Dark)"="darkblue","Down (Light)"="darkblue",
                                  "Up (Dark)"="darkred","Up (Light)"="darkred"), na.translate=FALSE) +
      scale_fill_manual(values=c("Down (Dark)"=alpha("darkblue",0.3),
                                 "Down (Light)"="white",
                                 "Up (Dark)"=alpha("darkred",0.3),
                                 "Up (Light)"="white"), na.translate=FALSE) +
      theme(legend.title=element_blank())
  }
  
  print(p.hold)
  invisible(p.hold)
}



```




## Males {.tabset}
### Individual genes

Loadings are identical to the full screen as we're using the same rotation matrix, they are printed here again for convenience. Rotation matrix was preserved so we can ensure we're validating the same signal. 


```{r, fig.width = 16,fig.height=6}
sex="M"

xx <- preProcess(df[df$Sex == sex  ,matched2], method = "YeoJohnson")
df_yj <- predict(xx, df[df$Sex == sex  ,matched2])

load("male_pca.rda")

res.full = read.csv(file="activity_males.csv")

res.full = res.full[res.full$Genotype=="dom",]

factoextra::fviz_eig(pca,ncp=15)

fviz_contrib(pca, choice = "var", axes = 1, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 2, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 3, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 4, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 5, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 6, top = 15)+coord_flip() + xlab("")


fit.p = list()
for(p in 1:6){    
    pc = as.data.frame(predict(pca, df_yj))
    pc$Genotype = df[df$Sex == sex  ,]$Genotype
    pc$file = factor(df[df$Sex == sex  ,]$file)
    pc$ID = factor(paste0(df[df$Sex == sex  ,]$Diet,df[df$Sex == sex  ,]$Replicate,df[df$Sex == sex  ,]$Genotype,df[df$Sex == sex  ,]$file))
    
    colnames(pc)[which(colnames(pc) == paste0("PC",p))] = "PCX"
    
    fit.p[[p]] = lme(PCX ~ Genotype, random=~1|file/ID,data=pc,weights = varIdent(form = ~1|Genotype),method="ML",control = lmeControl(msMaxIter = 500, msMaxEval = 500))
    
}


res.tmp=NULL
plots.dist = list()
plots.pc = list()
gns = unique(df$Genotype[df$Genotype != "WT"])
  
for(i in gns){
  print(i)
  for(f in 1:6){
      pc = as.data.frame(predict(pca, df_yj))
      pc$Genotype = df[df$Sex == sex  ,]$Genotype
      pc$file = factor(df[df$Sex == sex  ,]$file)
      plot.data=pc[ pc$Genotype %in% c("WT",i),]
      colnames(plot.data)[colnames(plot.data) == paste0("PC",f)] = "PCX"
      plot.data$Genotype = factor(plot.data$Genotype,levels=c("WT",i))
      
      fit = fit.p[[f]]
    
      g.ix = which(names(summary(fit)$coeff$fixed) == paste0("Genotype",i))
      ctrl.ixs = which(names(summary(fit)$coeff$fixed) != paste0("Genotype",i))[-1] # drop the intercept, we only want the non i genotype coeffs
          
      ctrl.contr = rep(-1/length(ctrl.ixs), length(ctrl.ixs))
      contrast = rep(1,length(summary(fit)$coeff$fixed))
      contrast[1] = 0 #We don't care about the WT mean aka the intercept
      contrast[ctrl.ixs] = ctrl.contr
      contrast[g.ix] = 1
      mean.c = abs(contrast) 
      mean.c[g.ix] = 0
      int = c(1,rep(0,length(mean.c)-1))
      v.WT = int
      v.WT[g.ix] = 1
      v.WT[1] = -1
      L <- list(v.WT,int)
    
      res = merge(confint(emmeans(fit, "Genotype") |>
        contrast(lapply(1,function(x) L[[x]]))),
        emmeans(fit, "Genotype") |>
        contrast(lapply(1,function(x) L[[x]])))
        
      mus = confint(emmeans(fit, "Genotype") |>
        contrast(lapply(2,function(x) L[[x]])))
      
      res$contrast = c(paste0(i, " vs WT"))
      
      plots.pc[[f]] = ggplot(plot.data,aes(x=Genotype,y=PCX,color = Genotype)) +
        geom_quasirandom()+theme_prism()+
        geom_boxplot(fill=NA,outliers = F)+ylab(paste0("PC",f)) +
        theme(strip.background = element_blank())+
        scale_color_manual(values=custom_colors )
                            
       if(res$p.value < 0.05){
            label = case_when(res$p.value<0.001 ~ "***",
            res$p.value<0.01 ~ "**",
            res$p.value<0.05 ~ "*")
            
            y <- max(plot.data$PCX)
            plots.pc[[f]] = plots.pc[[f]] + 
              annotate(geom="line",x = c(1,1.35),y = rep(y*1.15,2)) +
              annotate(geom="line",x = c(1.65,2),y = rep(y*1.15,2)) +
              annotate(geom="text",x=1.5,y=1.15*y,label=label,size=8,color="black",vjust=0.8)


       }
      res.tmp = rbind(res.tmp,data.frame(Genotype = i, p = paste0("PC",f),res))
  }
    # print(wrap_plots(plots.dist) + plot_layout(ncol=6, guides="collect" ))
    print(wrap_plots(plots.pc) + plot_layout(ncol=6, guides="collect" ))
}
    
res.tmp$Sex = "Male"

res.tmp$Result = case_when(res.tmp$p.value < 0.05 ~ "Difference",
                           TRUE~"Inconclusive") 


```


### Individual 
Must choose genotype(s)

```{r}

res.indv.m=NULL
#Add more genotypes here to get more individual analysis with rose plots
for(i in c("dom")){
  ix = which(colnames(df) %in% c("Genotype","file","Replicate"))
  
  for(p in names(matched2)){
    indv.mut = df[df$Sex == sex,c(matched2,ix)]
    indv.mut$ID = paste0(indv.mut$Genotype,indv.mut$file,indv.mut$Replicate)
    colnames(indv.mut)[1:51] = names(matched2)
    
    colnames(indv.mut)[colnames(indv.mut) == p] = "VarX"
    
    if(any(indv.mut$VarX<=0)){
      indv.mut$VarX = indv.mut$VarX + abs(min(indv.mut$VarX,na.rm=T)) + abs(min(indv.mut$VarX,na.rm=T))[abs(min(indv.mut$VarX,na.rm=T))>0]*0.1
    }
    
    fit = lme(log2(VarX) ~ Genotype, random= ~1|ID,data=indv.mut,weights = varIdent(form = ~1|Genotype),method="ML",control = lmeControl(msMaxIter = 500, msMaxEval = 500,opt = "optim"))
    
    g.ix = which(names(summary(fit)$coeff$fixed) == paste0("Genotype",i))
    ctrl.ixs = which(names(summary(fit)$coeff$fixed) != paste0("Genotype",i))[-1] # drop the intercept, we only want the non i genotype coeffs
    
    ctrl.contr = rep(-1/length(ctrl.ixs), length(ctrl.ixs))
    contrast = rep(1,length(summary(fit)$coeff$fixed))
    contrast[1] = 0 #We don't care about the WT mean aka the intercept
    contrast[ctrl.ixs] = ctrl.contr
    contrast[g.ix] = 1
    mean.c = abs(contrast)
    mean.c[g.ix] = 0
    int = c(1,rep(0,length(mean.c)-1))
    v.WT = int
    v.WT[g.ix] = 1
    v.WT[1] = -1
    L <- list(v.WT, int)
    
    
    res = merge(confint(emmeans(fit, "Genotype") |>
        contrast(lapply(1,function(x) L[[x]]))),
        emmeans(fit, "Genotype") |>
        contrast(lapply(1,function(x) L[[x]])))
        
    mus = confint(emmeans(fit, "Genotype") |>
                    contrast(lapply(2,function(x) L[[x]])))
    res$contrast = c(paste0(p, " vs WT"))
    
    
    
    res.indv.m=rbind(res.indv.m,data.frame(Genotype = i , Sex = sex, Outcome = p,res))
  }
}



write.csv(res.indv.m,file=paste0(file_path,"/",ifelse(MiMIC,"MiMIC","Trojan")  ,"individual_outcome_res_male.csv"))

pdf(paste0(ifelse(MiMIC,"MiMIC","Trojan"),"/bchs_rose_plots_combined_males_.pdf") , width=15, height=15)
rose_plot(res.indv.m, "M", "CSD", outcomes_24, mode="24h")
rose_plot(res.indv.m, "M", "CSD", outcomes_ld, mode="ld")
dev.off()

```



### Table


```{r}

DT::datatable(res.tmp)

write.csv(res.tmp,file=paste0(file_path,"/",ifelse(MiMIC,"MiMIC","Trojan") ,"male_activity_validation.csv"))


```


### Forest plot



```{r,fig.width = 5,fig.height=5.5}

res.tmp$PC = factor(res.tmp$p,levels=c("PC1","PC2","PC3","PC4","PC5","PC6"))

counts = aggregate(I(Result == "Difference") ~ Genotype,data=res.tmp,sum)
colnames(counts)[2] = "counts"
res.tmp = merge(res.tmp, counts,by="Genotype")

res.tmp$PC = factor(res.tmp$PC,levels=(unique(res.tmp$PC)))
res.tmp$counts = factor(res.tmp$counts)


ggplot(res.tmp, aes(y = estimate, x= p,color=Result)) + 
  geom_point(size=3) + geom_errorbar(aes(ymin = lower.CL,ymax=upper.CL),width=0.1) +      coord_flip() +theme_prism() + xlab("")+
  ylab("Mean difference") + ggtitle(i) + 
  geom_hline(yintercept = 0,linetype="dotted",color="grey",linewidth=1.5) + 
  scale_color_manual(values = c(Inconclusive = "black",Difference = "red"))



```



## Females {.tabset}
### Individual genes

Loadings are identical to the full screen as we're using the same rotation matrix, they are printed here again for convenience. Rotation matrix was preserved so we can ensure we're validating the same signal. 


```{r, fig.width = 16,fig.height=6}
sex="F"

xx <- preProcess(df[df$Sex == sex  ,matched2], method = "YeoJohnson")
df_yj <- predict(xx, df[df$Sex == sex  ,matched2])

load("female_pca.rda")

res.full = read.csv(file="activity_females.csv")


res.full = res.full[res.full$Genotype=="dom",]

factoextra::fviz_eig(pca,ncp=15)

fviz_contrib(pca, choice = "var", axes = 1, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 2, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 3, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 4, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 5, top = 15)+coord_flip() + xlab("")
fviz_contrib(pca, choice = "var", axes = 6, top = 15)+coord_flip() + xlab("")


fit.p = list()
for(p in 1:6){    
  pc = as.data.frame(predict(pca, df_yj))
  pc$Genotype = df[df$Sex == sex  ,]$Genotype
  pc$file = factor(df[df$Sex == sex  ,]$file)
  pc$ID = factor(paste0(df[df$Sex == sex  ,]$Diet,df[df$Sex == sex  ,]$Replicate,df[df$Sex == sex  ,]$Genotype,df[df$Sex == sex  ,]$file))
  
  colnames(pc)[which(colnames(pc) == paste0("PC",p))] = "PCX"
  
  fit.p[[p]] = lme(PCX ~ Genotype, random=~1|file/ID,data=pc,weights = varIdent(form = ~1|Genotype),method="ML",control = lmeControl(msMaxIter = 500, msMaxEval = 500))
  
}


res.tmp=NULL
plots.dist = list()
plots.pc = list()
gns = unique(df$Genotype[df$Genotype != "WT"])

for(i in gns){
  print(i)
  for(f in 1:6){
    pc = as.data.frame(predict(pca, df_yj))
    pc$Genotype = df[df$Sex == sex  ,]$Genotype
    pc$file = factor(df[df$Sex == sex  ,]$file)
    plot.data=pc[ pc$Genotype %in% c("WT",i),]
    colnames(plot.data)[colnames(plot.data) == paste0("PC",f)] = "PCX"
    plot.data$Genotype = factor(plot.data$Genotype,levels=c("WT",i))
    
    fit = fit.p[[f]]
    
    g.ix = which(names(summary(fit)$coeff$fixed) == paste0("Genotype",i))
    ctrl.ixs = which(names(summary(fit)$coeff$fixed) != paste0("Genotype",i))[-1] # drop the intercept, we only want the non i genotype coeffs
    
    ctrl.contr = rep(-1/length(ctrl.ixs), length(ctrl.ixs))
    contrast = rep(1,length(summary(fit)$coeff$fixed))
    contrast[1] = 0 #We don't care about the WT mean aka the intercept
    contrast[ctrl.ixs] = ctrl.contr
    contrast[g.ix] = 1
    mean.c = abs(contrast) 
    mean.c[g.ix] = 0
    int = c(1,rep(0,length(mean.c)-1))
    v.WT = int
    v.WT[g.ix] = 1
    v.WT[1] = -1
    L <- list(v.WT,int)
    
    res = merge(confint(emmeans(fit, "Genotype") |>
                          contrast(lapply(1,function(x) L[[x]]))),
                emmeans(fit, "Genotype") |>
                  contrast(lapply(1,function(x) L[[x]])))
    
    mus = confint(emmeans(fit, "Genotype") |>
                    contrast(lapply(2,function(x) L[[x]])))
    
    res$contrast = c(paste0(i, " vs WT"))
    
    plots.pc[[f]] = ggplot(plot.data,aes(x=Genotype,y=PCX,color = Genotype)) +
      geom_quasirandom()+theme_prism()+
      geom_boxplot(fill=NA,outliers = F)+ylab(paste0("PC",f)) +
      theme(strip.background = element_blank())+
      scale_color_manual(values=custom_colors )
    
    if(res$p.value < 0.05){
      label = case_when(res$p.value<0.001 ~ "***",
                        res$p.value<0.01 ~ "**",
                        res$p.value<0.05 ~ "*")
      
      y <- max(plot.data$PCX)
      plots.pc[[f]] = plots.pc[[f]] + 
        annotate(geom="line",x = c(1,1.35),y = rep(y*1.15,2)) +
        annotate(geom="line",x = c(1.65,2),y = rep(y*1.15,2)) +
        annotate(geom="text",x=1.5,y=1.15*y,label=label,size=8,color="black",vjust=0.8)
      
      
    }
    res.tmp = rbind(res.tmp,data.frame(Genotype = i, p = paste0("PC",f),res))
  }
  # print(wrap_plots(plots.dist) + plot_layout(ncol=6, guides="collect" ))
  print(wrap_plots(plots.pc) + plot_layout(ncol=6, guides="collect" ))
}

res.tmp$Sex = "Female"

res.tmp$Result = case_when(res.tmp$p.value < 0.05 ~ "Difference",
                           TRUE~"Inconclusive") 


```


### Individual 
Must choose genotype(s)

```{r}

res.indv.f=NULL
#Add more genotypes here to get more individual analysis with rose plots
for(i in c("dom")){
  ix = which(colnames(df) %in% c("Genotype","file","Replicate"))
  
  for(p in names(matched2)){
    indv.mut = df[df$Sex == sex,c(matched2,ix)]
    indv.mut$ID = paste0(indv.mut$Genotype,indv.mut$file,indv.mut$Replicate)
    colnames(indv.mut)[1:51] = names(matched2)
    
    colnames(indv.mut)[colnames(indv.mut) == p] = "VarX"
    
    if(any(indv.mut$VarX<=0)){
      indv.mut$VarX = indv.mut$VarX + abs(min(indv.mut$VarX,na.rm=T)) + abs(min(indv.mut$VarX,na.rm=T))[abs(min(indv.mut$VarX,na.rm=T))>0]*0.1
    }
    
    fit = lme(log2(VarX) ~ Genotype, random= ~1|ID,data=indv.mut,weights = varIdent(form = ~1|Genotype),method="ML",control = lmeControl(msMaxIter = 500, msMaxEval = 500,opt = "optim"))
    
    g.ix = which(names(summary(fit)$coeff$fixed) == paste0("Genotype",i))
    ctrl.ixs = which(names(summary(fit)$coeff$fixed) != paste0("Genotype",i))[-1] # drop the intercept, we only want the non i genotype coeffs
    
    ctrl.contr = rep(-1/length(ctrl.ixs), length(ctrl.ixs))
    contrast = rep(1,length(summary(fit)$coeff$fixed))
    contrast[1] = 0 #We don't care about the WT mean aka the intercept
    contrast[ctrl.ixs] = ctrl.contr
    contrast[g.ix] = 1
    mean.c = abs(contrast)
    mean.c[g.ix] = 0
    int = c(1,rep(0,length(mean.c)-1))
    v.WT = int
    v.WT[g.ix] = 1
    v.WT[1] = -1
    L <- list(v.WT, int)
    
    
    res = merge(confint(emmeans(fit, "Genotype") |>
                          contrast(lapply(1,function(x) L[[x]]))),
                emmeans(fit, "Genotype") |>
                  contrast(lapply(1,function(x) L[[x]])))
    
    mus = confint(emmeans(fit, "Genotype") |>
                    contrast(lapply(2,function(x) L[[x]])))
    res$contrast = c(paste0(p, " vs WT"))
    
    
    
    res.indv.f=rbind(res.indv.f,data.frame(Genotype = i , Sex = sex, Outcome = p,res))
  }
}



write.csv(res.indv.f,file=paste0(file_path,"/",ifelse(MiMIC,"MiMIC","Trojan")  ,"individual_outcome_res_female.csv"))

pdf(pdf(paste0(ifelse(MiMIC,"MiMIC","Trojan"),"/bchs_rose_plots_combined_females_.pdf") , width=15, height=15), width=15, height=15)
rose_plot(res.indv.f, "F", "CSD", outcomes_24, mode="24h")
rose_plot(res.indv.f, "F", "CSD", outcomes_ld, mode="ld")
dev.off()

```






### Table


```{r}

DT::datatable(res.tmp)

write.csv(res.tmp,file=paste0(file_path,"/",ifelse(MiMIC,"MiMIC","Trojan") ,"female_activity_validation.csv"))



```


### Forest plot



```{r,fig.width = 5,fig.height=5.5}

res.tmp$PC = factor(res.tmp$p,levels=c("PC1","PC2","PC3","PC4","PC5","PC6"))

counts = aggregate(I(Result == "Difference") ~ Genotype,data=res.tmp,sum)
colnames(counts)[2] = "counts"
res.tmp = merge(res.tmp, counts,by="Genotype")

res.tmp$PC = factor(res.tmp$PC,levels=(unique(res.tmp$PC)))
res.tmp$counts = factor(res.tmp$counts)


ggplot(res.tmp, aes(y = estimate, x= p,color=Result)) + 
  geom_point(size=3) + geom_errorbar(aes(ymin = lower.CL,ymax=upper.CL),width=0.1) +      coord_flip() +theme_prism() + xlab("")+
  ylab("Mean difference") + ggtitle(i) + 
  geom_hline(yintercept = 0,linetype="dotted",color="grey",linewidth=1.5) + 
  scale_color_manual(values = c(Inconclusive = "black",Difference = "red"))



```

